---
title: "McKeesport Parcel Lookup"
format:
  html:
    self-contained: true
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
---

```{css}
body {
  margin: 0 auto;
  padding: 0 1.5rem 3rem;
  font-family: system-ui, sans-serif;
}
h1 { margin-bottom: 0.25rem; }
.subtitle {
  color: #555;
  font-size: 0.95rem;
  margin-bottom: 1.25rem;
}
.map-wrapper {
  width: 100%;
  margin-bottom: 2rem;
}
/* Landscape map: constrained to body width, no horizontal scroll */
.leaflet-container {
  width:  100% !important;
  height: min(calc(100vw / 1.6), calc(100vh - 6rem)) !important;
  display: block;
  margin: 0 auto;
}
.notes-section {
  background: #f8f8f8;
  border-left: 4px solid #1a3a6b;
  padding: 1rem 1.5rem;
  border-radius: 0 4px 4px 0;
  font-size: 0.9rem;
  line-height: 1.6;
}
.notes-section h2 { margin-top: 0; font-size: 1.1rem; }
.notes-section h3 { font-size: 0.95rem; margin-top: 1.2rem; margin-bottom: 0.25rem; color: #1a3a6b; }
```

::: {.subtitle}
Interactive parcel lookup for McKeesport, PA. Hover over any parcel to see ownership type, owner name, use, assessed value, and sale history. Data: Allegheny County Assessor, October 2025.
:::

```{r setup}
library(sf)
library(dplyr)
library(stringr)
library(leaflet)
library(htmltools)

# ── Paths ──────────────────────────────────────────────────────────────────────
gdb_path  <- "McKee_Dev_Parcels.gdb"
gdb_layer <- "Parcels_Ownership_Dev"

# ── McKeesport center ──────────────────────────────────────────────────────────
center_lat <- 40.3425
center_lon <- -79.8491
```

```{r data}
# Read McKeesport parcels from GeoPackage (already filtered to McKeesport)
parcels <- st_read(gdb_path, layer = gdb_layer, quiet = TRUE) |>
  st_zm(drop = TRUE) |>
  st_transform(4326) |>
  mutate(across(where(is.character), str_trim)) |>
  mutate(across(where(\(x) inherits(x, "POSIXct") || inherits(x, "Date")),
                \(x) format(x, "%Y-%m-%d"))) |>
  mutate(Use_Type = if_else(Use_Type == "Med Desnity", "Med Density", Use_Type))
```

```{r labels}
# ── Helper formatters ──────────────────────────────────────────────────────────
fmt_dollar <- function(x) {
  ifelse(is.na(x) | x == 0, "\u2014",
    paste0("$", formatC(as.numeric(x), format = "f", digits = 0, big.mark = ",")))
}

fmt_date <- function(x) {
  ifelse(is.na(x) | str_trim(x) == "", "\u2014", str_trim(x))
}

# ── Ownership type classifier (translated from Own_Type_Final.py) ──────────────
classify_owner <- function(owner, taxdesc) {
  o <- toupper(coalesce(owner, ""))
  t <- toupper(coalesce(taxdesc, ""))

  # Rule 1: PURTA
  if (startsWith(t, "12 - PUBLIC UTILITY REALTY TAX")) return("Utilities (PURTA)")

  # Rule 2: Churches / religious — always Institution
  church_kw <- c("CHURCH", "PARISH", "PRESBYTERIAN", "BAPTIST", "CATHOLIC", "LUTHERAN",
                 "EPISCOPAL", "METHODIST", "TEMPLE", "SYNAGOGUE", "MOSQUE", "DIOCESE",
                 "MINISTR", "MISSION", "CONGREGATION")
  if (any(str_detect(o, fixed(church_kw)))) return("Institution")

  # Rule 3: City of McKeesport entities
  if (str_detect(o, fixed("CITY OF MCKEESPORT")) ||
      str_detect(o, fixed("CITY OF MC KEESPORT")) ||
      (str_detect(o, fixed("MCKEESPORT")) && str_detect(o, fixed("REDEVELOPMENT AUTHORITY"))) ||
      (str_detect(o, fixed("MCKEESPORT")) && str_detect(o, fixed("PARKING AUTHORITY"))) ||
      (str_detect(o, fixed("MCKEESPORT")) && str_detect(o, fixed("HOUSING AUTHORITY")))) {
    return("City")
  }

  # Rule 4: County entities
  if (str_detect(o, fixed("ALLEGHENY COUNTY")) || str_detect(o, fixed("COUNTY OF "))) {
    return("County")
  }

  # Rule 5: Tax-exempt → Institution
  if (t == "10 - EXEMPT") return("Institution")

  # Rule 6: False positive patterns (suppress institution keyword check)
  fp <- c("HOSPITALITY CORP", " LLC", " LP", " INC", "INVESTMENT", "RENTAL")
  has_fp <- any(str_detect(o, fixed(fp)))

  # Rule 7: Federal — Veterans Affairs → Institution; other Secretary → Other Public
  if (str_detect(o, fixed("SECRETARY OF VETERANS"))) return("Institution")
  if (str_detect(o, fixed("SECRETARY OF"))) return("Other Public")

  # Rule 8: Other nonprofits / charitable institutions (if no false positive)
  inst_kw <- c("FOUNDATION", "CHARITABLE", "COMMUNITY SERVICES",
               "MON YOUGH COMMUNITY SERVICES", "MON-YOUGH COMMUNITY SERVICES",
               "HOSPITAL", "YOUNG MENS CHRISTIAN", "YMCA", "YWCA",
               "SALVATION ARMY", "LODGE", "VETERANS", "WELFARE ASSOCIATION",
               "NONPROFIT", "NON-PROFIT", "NON PROFIT")
  if (!has_fp && any(str_detect(o, fixed(inst_kw)))) return("Institution")

  # Rule 9: Other public entities
  pub_kw <- c("AUTHORITY", "SCHOOL DISTRICT", "BOROUGH OF", "TOWNSHIP OF",
              "COMMONWEALTH OF", "UNITED STATES", "US GOVT", "U S GOVERNMENT",
              "PENNSYLVANIA TURNPIKE", "PORT AUTHORITY")
  if (any(str_detect(o, fixed(pub_kw)))) return("Other Public")

  # Rule 10: Private (fallback)
  "Private"
}

# Vectorize over all parcels
parcels <- parcels |>
  mutate(own_type = mapply(classify_owner, PROPERTYOWNER, TAXDESC))

# ── Color by Use_Type (development districts) ─────────────────────────────────
use_type_colors <- c(
  "Low Density"  = "#F5C518",
  "Med Density"  = "#F0A500",
  "High Density" = "#A89000",
  "Planned Unit" = "#E8650A",
  "Mixed Use"    = "#7BAFD4",
  "Hotel"        = "#4A7FBF",
  "CSA"          = "#2E6B1A"
)

# ── Color by ownership type ────────────────────────────────────────────────────
own_colors <- c(
  "Private"           = "#4a90d9",  # blue
  "City"              = "#e8890c",  # orange
  "County"            = "#7b1fa2",  # purple
  "Institution"       = "#2e7d32",  # green
  "Other Public"      = "#b0bec5",  # light slate
  "Utilities (PURTA)" = "#78909c"   # slate
)

# ── Build hover label HTML ─────────────────────────────────────────────────────
parcels <- parcels |>
  mutate(
    fill_color      = unname(coalesce(own_colors[own_type], "#78909c")),
    dist_fill_color = unname(coalesce(use_type_colors[Use_Type], "#cccccc")),
    addr_str     = str_trim(paste(
      coalesce(as.character(PROPERTYHOUSENUM), ""),
      coalesce(PROPERTYADDRESS, "")
    )),
    dist_html = ifelse(
      !is.na(Dist_num) & Dist_num > 0,
      paste0(
        "<hr style='margin:6px 0;border:none;border-top:1px solid #eee'>",
        "<b>Development District</b><br>",
        "District #: <b>", Dist_num, "</b><br>",
        "Use type: ", coalesce(Use_Type, "\u2014")
      ),
      ""
    ),
    label_html = paste0(
      "<div style='font-family:system-ui,sans-serif;font-size:12px;line-height:1.5;",
      "max-width:300px;word-wrap:break-word;overflow-wrap:break-word'>",
      "<b style='font-size:14px'>", addr_str, "</b><br>",
      "<span style='color:#888;font-size:11px'>PIN: ", PIN, "</span>",

      "<hr style='margin:6px 0;border:none;border-top:1px solid #eee'>",
      "<b>Owner</b><br>",
      coalesce(PROPERTYOWNER, "\u2014"), "<br>",
      "Ownership type: <b>", own_type, "</b><br>",
      "Tax status: ", coalesce(TAXDESC, "\u2014"),

      "<hr style='margin:6px 0;border:none;border-top:1px solid #eee'>",
      "<b>Classification</b><br>",
      "Class: ", coalesce(CLASSDESC, "\u2014"), "<br>",
      "Use: ", coalesce(USEDESC, "\u2014"),

      "<hr style='margin:6px 0;border:none;border-top:1px solid #eee'>",
      "<b>Assessed Value</b><br>",
      "County total: ", fmt_dollar(COUNTYTOTAL), "<br>",
      "Fair market: <b>", fmt_dollar(FAIRMARKETTOTAL), "</b><br>",
      "Land: ", fmt_dollar(FAIRMARKETLAND), "<br>",
      "Building: ", fmt_dollar(FAIRMARKETBUILDING),

      "<hr style='margin:6px 0;border:none;border-top:1px solid #eee'>",
      "<b>Sale History</b><br>",
      "Last sale: ", fmt_date(SALEDATE), " &mdash; ", fmt_dollar(SALEPRICE),
      ifelse(!is.na(SALEDESC) & SALEDESC != "", paste0(" <i>(", SALEDESC, ")</i>"), ""), "<br>",
      "Prior: ", fmt_date(PREVSALEDATE), " &mdash; ", fmt_dollar(PREVSALEPRICE), "<br>",
      "2nd prior: ", fmt_date(PREVSALEDATE2), " &mdash; ", fmt_dollar(PREVSALEPRICE2),

      dist_html,

      "</div>"
    )
  )
```

```{r export}
library(openxlsx)

# ── Select and rename columns for export ──────────────────────────────────────
col_map <- c(
  # Unique IDs
  "PIN"                    = "PIN",
  "MAPBLOCKLOT"            = "Map Block Lot",
  # Development district (moved to front)
  "Dist_num"               = "District Number",
  "Use_Type"               = "New Use Type",
  # Property address
  "PROPERTYHOUSENUM"       = "House Number",
  "PROPERTYADDRESS"        = "Property Address",
  "MUNICODE"               = "Muni Code",
  "MUNIDESC"               = "Municipality",
  "NEIGHDESC"              = "Neighborhood",
  # Owner
  "PROPERTYOWNER"          = "Property Owner",
  "own_type"               = "Owner Type",
  "CHANGENOTICEADDRESS1"   = "Owner Mailing Address 1",
  "CHANGENOTICEADDRESS2"   = "Owner Mailing Address 2",
  "CHANGENOTICEADDRESS3"   = "Owner Mailing Address 3",
  "CHANGENOTICEADDRESS4"   = "Owner Mailing Address 4",
  # Tax / use
  "TAXCODE"                = "Tax Code",
  "TAXDESC"                = "Tax Status",
  "CLASS"                  = "Class",
  "CLASSDESC"              = "Class Description",
  "USECODE"                = "Use Code",
  "USEDESC"                = "Use Description",
  "HOMESTEADFLAG"          = "Homestead Flag",
  # Parcel value
  "COUNTYLAND"             = "County Land Value",
  "COUNTYBUILDING"         = "County Building Value",
  "COUNTYTOTAL"            = "County Total Value",
  "FAIRMARKETLAND"         = "Fair Market Land",
  "FAIRMARKETBUILDING"     = "Fair Market Building",
  "FAIRMARKETTOTAL"        = "Fair Market Total",
  # Sale history
  "SALEDATE"               = "Sale Date",
  "SALEPRICE"              = "Sale Price",
  "SALECODE"               = "Sale Code",
  "SALEDESC"               = "Sale Description",
  "PREVSALEDATE"           = "Prior Sale Date",
  "PREVSALEPRICE"          = "Prior Sale Price",
  "PREVSALEDATE2"          = "2nd Prior Sale Date",
  "PREVSALEPRICE2"         = "2nd Prior Sale Price",
  # Physical
  "STYLEDESC"              = "Style",
  "YEARBLT"                = "Year Built",
  "CONDITIONDESC"          = "Condition",
  "GRADEDESC"              = "Grade",
  "FINISHEDLIVINGAREA"     = "Finished Living Area (Sq Ft)",
  "LOTAREA"                = "Lot Area (Sq Ft)"
)

# Only keep source columns that exist in the data
valid_src <- intersect(names(col_map), names(parcels))
valid_map <- col_map[valid_src]

export_df <- st_drop_geometry(parcels)[, valid_src] |>
  rename(!!!setNames(valid_src, valid_map))

export_dist <- export_df |> filter(!is.na(`District Number`) & `District Number` > 0)

dir.create("output", showWarnings = FALSE)

wb <- createWorkbook()

addWorksheet(wb, "Dev District Parcels")
writeDataTable(wb, "Dev District Parcels", export_dist,
               tableStyle = "TableStyleMedium2", withFilter = TRUE)

addWorksheet(wb, "All Parcels")
writeDataTable(wb, "All Parcels", export_df,
               tableStyle = "TableStyleMedium9", withFilter = TRUE)

setColWidths(wb, "Dev District Parcels", cols = seq_along(valid_map), widths = "auto")
setColWidths(wb, "All Parcels",          cols = seq_along(valid_map), widths = "auto")

saveWorkbook(wb, "output/mckeesport_parcels.xlsx", overwrite = TRUE)
```

```{r map}
# ── Development district subsets ────────────────────────────────────────────
dist_parcels <- parcels |> filter(!is.na(Dist_num) & Dist_num > 0)

dist_labels <- dist_parcels |>
  select(Dist_num) |>
  group_by(Dist_num) |>
  summarise(.groups = "drop")

dist_coords <- st_coordinates(st_centroid(dist_labels))
dist_labels$lng <- dist_coords[, 1]
dist_labels$lat <- dist_coords[, 2]

m <- leaflet(
  options = leafletOptions(
    minZoom   = 13,
    maxZoom   = 19,
    maxBounds = list(
      list(center_lat - 0.06, center_lon - 0.09),
      list(center_lat + 0.06, center_lon + 0.09)
    )
  )
) |>

  # ── Basemaps ────────────────────────────────────────────────────────────────
  addProviderTiles("CartoDB.Positron",  group = "Light (default)") |>
  addProviderTiles("Esri.WorldImagery", group = "Satellite") |>

  # ── Base parcel layer (grey, always on, carries tooltips + popups) ───────────
  addPolygons(
    data        = parcels,
    fillColor   = "#cccccc",
    fillOpacity = 0.4,
    color       = "#555",
    weight      = 0.7,
    opacity     = 0.8,
    label        = ~lapply(label_html, HTML),
    labelOptions = labelOptions(
      style     = list("font-size" = "12px", "max-width" = "300px"),
      sticky    = FALSE,
      direction = "auto"
    ),
    popup        = ~label_html,
    popupOptions = popupOptions(maxWidth = 320, closeOnClick = FALSE),
    group = "Parcels"
  ) |>

  # ── Ownership type color layer (toggleable, hidden by default) ───────────────
  addPolygons(
    data        = parcels,
    fillColor   = ~fill_color,
    fillOpacity = 0.55,
    color       = "#333",
    weight      = 0.7,
    opacity     = 0.8,
    label        = ~lapply(label_html, HTML),
    labelOptions = labelOptions(
      style     = list("font-size" = "12px", "max-width" = "300px"),
      sticky    = FALSE,
      direction = "auto"
    ),
    popup        = ~label_html,
    popupOptions = popupOptions(maxWidth = 320, closeOnClick = FALSE),
    group = "Ownership Type"
  ) |>

  # ── Development districts: colored parcels ───────────────────────────────────
  addPolygons(
    data        = dist_parcels,
    fillColor   = ~dist_fill_color,
    fillOpacity = 0.65,
    color       = "#555",
    weight      = 0.7,
    opacity     = 0.8,
    label        = ~lapply(label_html, HTML),
    labelOptions = labelOptions(
      style     = list("font-size" = "12px", "max-width" = "300px"),
      sticky    = FALSE,
      direction = "auto"
    ),
    popup        = ~label_html,
    popupOptions = popupOptions(maxWidth = 320, closeOnClick = FALSE),
    group = "Development Districts"
  ) |>
  addLabelOnlyMarkers(
    data         = dist_labels,
    lng          = ~lng,
    lat          = ~lat,
    label        = ~as.character(Dist_num),
    labelOptions = labelOptions(
      noHide    = TRUE,
      direction = "center",
      textOnly  = TRUE,
      style     = list(
        "font-weight"   = "bold",
        "font-size"     = "12px",
        "color"         = "#333",
        "background"    = "rgba(255,255,255,0.75)",
        "border"        = "none",
        "padding"       = "2px 4px",
        "border-radius" = "3px"
      )
    ),
    group = "Development Districts"
  ) |>

  # ── Legend: Ownership Type ──────────────────────────────────────────────────
  addLegend(
    position = "bottomright",
    colors   = c("#4a90d9", "#e8890c", "#7b1fa2", "#2e7d32", "#b0bec5", "#78909c"),
    labels   = c("Private", "City", "County", "Institution", "Other Public",
                 "Utilities (PURTA)"),
    title    = "Ownership Type<br><small>faded = tax-exempt</small>",
    opacity  = 0.85,
    layerId  = "legend-ownership",
    group    = "Ownership Type"
  ) |>

  # ── Legend: Development Districts ──────────────────────────────────────────
  addLegend(
    position = "bottomright",
    colors   = c("#F5C518", "#F0A500", "#A89000", "#E8650A", "#7BAFD4", "#4A7FBF", "#2E6B1A"),
    labels   = c("Low Density", "Med Density", "High Density", "Planned Unit",
                 "Mixed Use", "Hotel", "CSA"),
    title    = "Development District Use",
    opacity  = 0.85,
    layerId  = "legend-districts",
    group    = "Development Districts"
  ) |>

  # ── Layer control ───────────────────────────────────────────────────────────
  addLayersControl(
    baseGroups    = c("Light (default)", "Satellite"),
    overlayGroups = c("Parcels", "Ownership Type", "Development Districts"),
    position      = "topright",
    options       = layersControlOptions(collapsed = FALSE)
  ) |>

  hideGroup("Ownership Type") |>
  hideGroup("Development Districts") |>

  addScaleBar(position = "bottomleft") |>

  fitBounds(
    lng1 = center_lon - 0.03, lat1 = center_lat - 0.02,
    lng2 = center_lon + 0.03, lat2 = center_lat + 0.02
  ) |>

  # ── Legend toggle + layer control section labels ─────────────────────────────
  htmlwidgets::onRender("
    function(el, x) {
      var map = this;

      function legendById(layerId) {
        return el.querySelector('.leaflet-bottom.leaflet-right [data-layerid=\"' + layerId + '\"]') ||
               el.querySelector('.leaflet-bottom.leaflet-right .info.legend');
      }

      // Track legend visibility by layerId
      var legendIds = {
        'Ownership Type': 'legend-ownership',
        'Development Districts': 'legend-districts'
      };

      // Hide both legends on load (layers start hidden)
      Object.values(legendIds).forEach(function(lid) {
        var el2 = document.querySelector('[data-layerid=\"' + lid + '\"]');
        if (el2) el2.style.display = 'none';
      });

      map.on('overlayremove', function(e) {
        var lid = legendIds[e.name];
        if (lid) {
          var el2 = document.querySelector('[data-layerid=\"' + lid + '\"]');
          if (el2) el2.style.display = 'none';
        }
      });

      map.on('overlayadd', function(e) {
        var lid = legendIds[e.name];
        if (lid) {
          var el2 = document.querySelector('[data-layerid=\"' + lid + '\"]');
          if (el2) el2.style.display = '';
        }
      });

      // Inject section labels into the layer control
      var ctrl = el.querySelector('.leaflet-control-layers');
      if (ctrl) {
        var baseDiv = ctrl.querySelector('.leaflet-control-layers-base');
        if (baseDiv) {
          var baseLabel = document.createElement('div');
          baseLabel.style.cssText = 'font-weight:600;font-size:11px;color:#555;text-transform:uppercase;letter-spacing:0.04em;margin-bottom:4px;padding-top:2px';
          baseLabel.textContent = 'Basemap Style';
          baseDiv.insertBefore(baseLabel, baseDiv.firstChild);
        }
        var overlayDiv = ctrl.querySelector('.leaflet-control-layers-overlays');
        if (overlayDiv) {
          var overlayLabel = document.createElement('div');
          overlayLabel.style.cssText = 'font-weight:600;font-size:11px;color:#555;text-transform:uppercase;letter-spacing:0.04em;margin-bottom:4px;margin-top:6px;padding-top:4px;border-top:1px solid #ddd';
          overlayLabel.textContent = 'Layer Data';
          overlayDiv.insertBefore(overlayLabel, overlayDiv.firstChild);
        }
      }
    }
  ")

div(class = "map-wrapper", m)
```

---

::: {.notes-section}

## Data Notes

### Source & Currency
Parcel boundaries from the Allegheny County Division of Computer Services GIS Group (October 2025). Assessment attributes from the Allegheny County Assessor file, also October 2025. Both datasets reflect the 2025 tax year assessment.

### Field Definitions

**Ownership Type** — Derived by applying keyword and tax-status rules to the property owner name and tax description:
- **Private** — private individuals, LLCs, corporations, and partnerships
- **City** — City of McKeesport and its authorities (Redevelopment, Housing, Parking)
- **County** — Allegheny County entities
- **Institution** — churches, nonprofits, hospitals, foundations, and other tax-exempt charitable organizations
- **Other Public** — school districts, other municipalities, state/federal agencies, and other authorities
- **Utilities (PURTA)** — public utility infrastructure subject to the Public Utility Realty Tax Act

**Tax Status** — Taxable (T), Exempt (E), or PURTA (P). Exempt parcels are rendered at lower opacity on the map.

**Fair Market Value** — The assessor's estimate of open-market value. Land and building components are shown separately.

**County Total** — The assessed value used for county millage calculations. In Allegheny County, assessed value is set at a fixed percentage of fair market value based on a base year.

**Sale History** — Up to three prior sale records: date, price, and sale type. A $0 price or "—" typically indicates a deed transfer, estate transfer, or other non-arm's-length transaction.

### How to Use
Hover over any parcel to see a detailed label. Zoom in for individual lot detail. Use the layer control (top right) to toggle between basemaps. The scale bar (bottom left) adjusts as you zoom.

:::
